---
title: "Machine Learning final project"
author: "Han Zhang"
date: "August 18, 2017"
output: 
  html_document: 
    keep_md: yes
---

## R Markdown

The goal of this project is to predict the manner in which human did the exercise. This is the "classe" variable in the training set. I built a random forest model to predict the "classe". I will also use the prediction model to predict 20 different test cases.


## Reading data and cleaning

```{r}
training_Url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testing_Url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train <- read.csv(training_Url, header = T, na.strings = c("NA", "", "DIV/0!"))
test <- read.csv(testing_Url, header = T, na.strings = c("NA", "", "DIV/0!"))

# Summarize raw data
dim(train)
dim(test)

# cleaning
train_1 <- train[,!apply(train, 2, function(x) {sum(is.na(x))>2500})]
test_1 <- test[,!apply(test, 2, function(x) {sum(is.na(x))>10})]
training <- train_1[, c(8:dim(train_1)[2])]
testing <- test_1[, c(8:dim(test_1)[2])]
summary(training)
```


## Data preprocessing

```{r}
library(caret)
library(rattle)
library(randomForest)
library(survival)

preobj <- preProcess(training[, 1:52], method = c("center", "scale", "knnImpute"))
newtrain <- predict(preobj, training[, c(1:52)])
newtrain$classe <- training$classe
newtest <- predict(preobj, testing[, c(1:52)])
newtest$problem_id <- testing$problem_id

inTrain <- createDataPartition(y = newtrain$classe, p=0.7, list = F)
train_Train <- newtrain[inTrain,]
train_Test <- newtrain[-inTrain,]
```


## Random Forest Model

# Since Random Forest model is great for dealing with category prediction questions, here I use it first to build a model. 70% of training data was used to build the model and cross validation was used in the model to avoid overfitting. The model was tested on another 30% training data to check the performance.

```{r}
set.seed(12345)
modelfit1 <- train(classe~., data=train_Train, method="rf", trControl=trainControl(method="repeatedcv", number=5, repeats=3))
plot(varImp(modelfit1), top=20,  main="The most importance 20 variables")
```
# Data validation on RF Model

```{r}
pred_1 <- predict(modelfit1, newdata=train_Test[, 1:52])
confusionMatrix(pred_1, train_Test$classe)
```
The model showed great accuracy (99.2%), so the out of sample error = 1 - accuracy = 0.8%. Since the random forest model give great performance, so I will use this model to predict the testing dataset.


## Apply model1 to validation dataset

```{r}
prediction <- predict(modelfit1, newdata=newtest[, 1:52])
```

## Conclusion

The ConfusionMatrix achieved more than 99% sensitivity and specificity  and only 0.8% out of sample error. This model was great for the final prediction The prediction for validation data No.1-20 were:

```{r}
prediction
```
